---
title: "Blue-talk — Project Structure"
subtitle: "Decentralized mesh messaging network in Go"
author: "Eddy Odero"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Blue-talk is a decentralized, multi-hop mesh messaging and file-transfer
network. Devices ("nodes") discover each other and exchange encrypted
messages and files over Bluetooth and local Wi-Fi, without needing an
internet connection or a central server.

The project doubles as a structured way to learn Go: rather than studying
networking, concurrency, and cryptography in isolation, each concept is
introduced at the point the project actually needs it.

## Repository layout

```{bash, eval=FALSE}
blue-talk/
├── go.mod
├── README.md
├── PROJECT_STRUCTURE.Rmd
├── cmd/
│   └── node/
│       └── main.go          # entrypoint: starts listener + optional dial
├── internal/
│   ├── node/
│   │   └── node.go          # Node struct, ID, listen/dial logic
│   ├── message/
│   │   └── message.go       # Message struct + encode/decode
│   ├── discovery/           # (later) UDP broadcast / mDNS
│   ├── routing/             # (later) routing table + algorithms
│   ├── transport/           # (later) Bluetooth / Wi-Fi transports
│   ├── crypto/              # (later) TLS + public-key encryption
│   └── storage/             # (later) SQLite persistence
└── web/                      # frontend (tech stack TBD)
```

## Tech stack

| Layer            | Technology                      | Why                                  |
|-------------------|----------------------------------|----------------------------------------|
| Core language     | Go                               | Networking, concurrency, servers       |
| Node runtime      | Go application                   | Every device becomes a node            |
| Messaging         | TCP (initial)                    | Easy to understand and debug           |
| Discovery         | UDP broadcast / mDNS (initial)   | Find nearby nodes                      |
| Mesh routing      | Custom routing protocol          | Main learning/project challenge        |
| Bluetooth         | BLE                              | Nearby-device communication            |
| Local networking  | Wi-Fi / Wi-Fi Direct              | Faster communication                   |
| File transfer     | TCP/HTTP (initial), chunked      | Reliable large transfers               |
| Encryption        | TLS + public-key cryptography     | Secure end-to-end messages             |
| Storage           | SQLite                           | Messages, devices, routes              |
| Server            | Ubuntu Server                     | Optional supernode                     |
| API               | Go HTTP API                       | Communication between UI and node      |
| Frontend          | *TBD*                             | See "Frontend decision" below          |
| Containers        | Docker                            | Reproducible deployment, node simulation |
| Testing           | Go tests + network simulation     | Test mesh behavior                     |
| Version control   | Git + GitHub                      | Source control                         |

## Frontend decision (open)

Two candidate directions, not yet finalized:

1. **Web frontend** — React + TypeScript (+ Vite), served by the Go API.
   Familiar tooling; still HTML/CSS/DOM under the hood, just abstracted
   away by the framework.
2. **Native desktop app** — Fyne (a pure-Go GUI toolkit). No HTML/CSS/JS
   anywhere; the UI is written entirely in Go. A new toolkit to learn, but
   keeps the whole project in one language.

Target layout either way: an Outlook-style three-pane view — node/peer list,
conversation list, message detail pane.

## Message protocol

```{json, eval=FALSE}
{
  "type": "message",
  "id": "abc123",
  "source": "nodeA",
  "destination": "nodeE",
  "ttl": 8,
  "payload": "Hello!"
}
```

Key fields:

- **id** — unique message ID, so a node can recognize and drop duplicates it has already forwarded
- **ttl** — time to live, so a message can't circulate the mesh forever

## Development roadmap (10 stages)

```{r roadmap, echo=FALSE}
stages <- data.frame(
  Stage = 1:10,
  Focus = c(
    "Go networking fundamentals",
    "Two-node TCP messaging",
    "Automatic discovery (UDP broadcast / mDNS)",
    "Three+ node network",
    "Routing (flooding -> smarter algorithms)",
    "Store & forward for offline peers",
    "Chunked file transfer",
    "End-to-end encryption",
    "Bluetooth / Wi-Fi transport integration",
    "Ubuntu supernode + Docker packaging"
  )
)
knitr::kable(stages)
```

## Current status

Working through Stage 1 + 2: a two-node TCP handshake exchanging a shared
JSON `Message` struct. No discovery, routing, Bluetooth, or encryption yet —
deliberately kept minimal until the basic node-to-node link works.

## Development environment

- Windows PC: VS Code, Go, Git, Docker Desktop, WSL2/Ubuntu
- Ubuntu Server: reserved for the eventual supernode role
- No special hardware required for the current stage — everything runs over local TCP/UDP
